---
title: "p8105_hw2_jz3425"
author: "Jibei Zheng"
date: 10/2/2021
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Reading Mr. Trash Wheel dataset.

```{r, warning = FALSE}
#read Mr. Trash Wheel data set and clean
mr_trash_wheel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls)) 

mr_trash_wheel_df
```

Reading precipitation data for 2018 and 2019.

```{r}
#read precipitation data and create year variable
pre_18_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  relocate(year)

pre_19_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019) %>% 
  relocate(year)

#combine 18 and 19 dataset and convert month to a character variable
precipitation_df = 
  bind_rows(pre_18_df, pre_19_df) %>% 
  mutate(month = month.name[month])

precipitation_df
```

The Mr. Trash Wheel dataset includes `r nrow(mr_trash_wheel_df)` dumpsters and `r ncol(mr_trash_wheel_df)` variables from `r min(pull(mr_trash_wheel_df, year))`-`r max(pull(mr_trash_wheel_df, year))`. Data include `r names(mr_trash_wheel_df[5:6])` of trash and categories of trash: `r names(mr_trash_wheel_df[7:14])`.

The precipitation dataset lists the precipitation for each month in 2018 and 2019. The data of last 6 months of 2019 is missing. 

The total precipitation in 2018 is `r precipitation_df[13,3]`.

The median number of sports balls in a dumpster in 2019 is `r median(pull(mr_trash_wheel_df, sports_balls))`.

## Problem 2

Reading pols-month dataset.

```{r, message = FALSE}
#read pols_month data and separate the date variable
pols_month_df = 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day"), "-") %>% 
  mutate(month = as.numeric(month)) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(year = as.numeric(year))

#create president variable
pres = factor(c("dem", "gop"))
pols_month_df = 
  mutate(pols_month_df, president = pres[prez_gop + 1]) %>% 
  select(-prez_gop) %>% 
  select(-prez_dem) %>% 
  select(-day)

pols_month_df
```

Reading snp dataset.

```{r, message = FALSE}
#read and clean snp data
snp_df = 
  read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  separate(date, into = c("year", "month", "day"), "/") %>% 
  mutate(year = as.numeric(year) + 2000) %>% 
  mutate(month = as.numeric(month)) %>% 
  arrange(year, month) %>% 
  mutate(month = month.name[month]) %>% 
  select(-day)

snp_df
```

Reading unemployment dataset.

```{r, message = FALSE}
#read and clean unemployment dadta
unemp_df = 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv", col_names = c("year", month.name), skip = 1) %>% 
  pivot_longer(
    January:December,
    names_to = "month",
    values_to = "unemployment"
    ) %>% 
  mutate(year = as.numeric(year))

unemp_df
```

Join the datasets.

```{r, message = FALSE}
#join plos_month_df and snp_df
temp = left_join(pols_month_df, snp_df)

#join unemp_df into it
final_df = left_join(temp, unemp_df)

final_df
```

The "pols_month_df" dataset contains `r nrow(pols_month_df)` observations of `r ncol(pols_month_df)` variables related to the number of national politicians who are democratic or republican at any given time from `r min(pull(pols_month_df, year))`-`r max(pull(pols_month_df, year))`.

The "snp_df" dataset contains `r nrow(snp_df)` observatiosn of `r ncol(snp_df)` variables related to Standard & Poor’s stock market index from `r min(pull(snp_df, year))`-`r max(pull(snp_df, year))`.

The "unemp_df" dataset contains `r nrow(unemp_df)` observations of `r ncol(unemp_df)` variables related to the percentage of unemployment in each month from `r min(pull(unemp_df, year))`-`r max(pull(unemp_df, year))`.

The "final_df" dataset joins these three datasets and contains `r nrow(final_df)` observations of `r ncol(final_df)` variables including president, snp index, percentage of unemployment, from `r min(pull(final_df, year))`-`r max(pull(final_df, year))`.

## Problem 3

Reading the popular baby names dataset.

```{r, message = FALSE}
#read and clean popular baby names dataset
names_df =
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(gender = str_to_lower(gender)) %>% 
  mutate(ethnicity = str_to_lower(ethnicity)) %>% 
  mutate(childs_first_name = str_to_lower(childs_first_name)) %>% 
  distinct()

#create a table showing the rank in popularity of the name “Olivia” over time
olivia_df = 
  filter(names_df, childs_first_name == "olivia") %>% 
  arrange(year_of_birth) %>% 
  select(ethnicity, year_of_birth, rank) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "rank"
  )

olivia_df

#create a table showing the most popular name among male children over time
most_male_df= 
  filter(names_df, gender == "male", rank == 1) %>% 
  arrange(year_of_birth) %>% 
  select(-gender, -rank, -count) %>% 
  pivot_wider(
    names_from = "year_of_birth",
    values_from = "childs_first_name"
  )

most_male_df
```

