---
title: "p8105_hw2_jz3425"
author: "Jibei Zheng"
date: 10/2/2021
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Reading Mr. Trash Wheel dataset.

```{r, warning = FALSE}
mr_trash_wheel_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  drop_na(dumpster) %>% 
  mutate(sports_balls = round(sports_balls)) 

mr_trash_wheel_df
```

Reading precipitation data for 2018 and 2019.

```{r}
pre_18_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018) %>% 
  relocate(year)

pre_19_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019) %>% 
  relocate(year)

precipitation_df = 
  bind_rows(pre_18_df, pre_19_df) %>% 
  mutate(month = c(month.name, NA, month.name, NA))

precipitation_df
```

The Mr. Trash Wheel dataset includes data of `r nrow(mr_trash_wheel_df)` dumpsters from `r min(pull(mr_trash_wheel_df, year))`-`r max(pull(mr_trash_wheel_df, year))`. Data include `r names(mr_trash_wheel_df[5:6])` of trash and categories of trash: `r names(mr_trash_wheel_df[7:14])`.

The precipitation dataset lists the precipitation for each month in 2018 and 2019. The data of last 6 months of 2019 is missing. 

The total precipitation in 2018 is `r precipitation_df[13,3]`.

The median number of sports balls in a dumpster in 2019 is `r median(pull(mr_trash_wheel_df, sports_balls))`.